---
title: "Mining biosynthetic gene clusters in Carnobacterium maltaromaticum by interference competition network and genome analysis"
author: "F. Borges"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    keep_md: yes
    number_sections: yes
    toc : yes
    toc_float : yes
  word_document: 
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE) # to hide the code in the final report
knitr::opts_chunk$set(message = FALSE) # to hide the messages given in the output
knitr::opts_chunk$set(warning = FALSE) # to hide the warnings
knitr::opts_chunk$set(dpi = 300)
#knitr::opts_chunk$set(cache=TRUE) #les code chunk, excepté s’ils sont modifiés, ne seront évalués qu’une seule fois
```

# NOTE
The compilation of this Rmarkdown file will produce the graphics as separated files. See in the directory "output".  

# LIBRARIES
```{r}
library(tidyverse)
library(ComplexHeatmap)
library(viridis)
library(ggrepel)
library(ggthemes)
library(ggprism)
library(factoextra)
```


# DATA PREPARATION
Seed for reproducible code
```{r}
set.seed(666)
```

Load the names of the strains described in (Ramia et al, 2020)
```{r}
strains_ramia <- read.csv2(here::here("data-clean","strains_ramia2020.csv"),
                check.names = F, 
                dec = ",",
                stringsAsFactors = T)

```

Load the graph data
```{r}
mat <- read.csv2(here::here("data-clean","20220201_INHIB_MAT_ROW-SEND_clean2.csv"),
                check.names = F, 
                dec = ",",
                stringsAsFactors = T)

names(mat)[1] <- "from"
```

load the table indicating which strains have been genome sequenced
```{r}
genome <- read.csv2(here::here("data-clean","genome_bin.csv"),
                check.names = F, 
                dec = ",",
                stringsAsFactors = T)

genome$genome <- as.character(genome$genome)
```

```{r}
genome_yes <- genome %>% 
  filter(genome==1)
genome_yes <- genome_yes$strain
genome_yes <- droplevels(genome_yes)

```

How many genomes were analyzed ?
```{r}
length(genome_yes)
```


Define the strains inhibiting EGDe by the kmeans clustering method
```{r}
# make dataframe with the EGDe inhibition values
EGDe_inhib <- mat %>% 
  select(from, EGDe) %>% 
  column_to_rownames("from")
# Define the number of clusters 
EGDe_inhib_sc <- scale(EGDe_inhib)
fviz_nbclust(EGDe_inhib_sc, kmeans, method = "silhouette")+
  labs(subtitle = "Silhouette method") 
# => 2 clusters
# Kmeans clustering with the value of 2 clusters
km.out = kmeans(EGDe_inhib_sc,centers=2,nstart =20) 
EGDe_inhib <-  data.frame(km.out$cluster)
EGDe_inhib <- rownames_to_column(EGDe_inhib, "from")
colnames(EGDe_inhib)[2] <- "EGDe_bin"
EGDe_inhib <- EGDe_inhib %>% 
  mutate(EGDe_bin=ifelse(EGDe_bin==2, 1,0))
```


Build a binary matrix with all the data by using 300 as a threshold for Carno Vs Carno. For Carno Vs Listeria => use the EGDe_bin dataframe defined above
```{r}
mat_bin <- mat

for (i in 2:77) {   
  mat_bin[ , i]=ifelse(mat[ , i]>300, 1, 0)
}

mat_bin <- full_join(mat_bin,EGDe_inhib,by="from")


row.names(mat_bin) <- mat_bin$from
mat_bin <- select(mat_bin,-1)
```
Binary matrix without the values for Listeria
```{r}
mat_df_bin <- select(mat_bin, 1:76)
```

Calculation of the degrees
```{r}
# Degrees
## creation of an empty dataframe to store the results of calculation we'll do after
deg_receiver <- data.frame(strain=colnames(mat_df_bin),
                           Receiver=rep(0, ncol(mat_df_bin)))

deg_sender <- data.frame(strain=rownames(mat_df_bin),
                         Sender=rep(0, nrow(mat_df_bin)))


## Calculation of the Receiver degrees
for (i in 1:nrow(mat_df_bin)){
  deg_receiver[i, 2]<-sum(mat_df_bin[ ,i])
}

## Calculation of the Sender degrees
for (i in 1:ncol(mat_df_bin)){
  deg_sender[i, 2]<-sum(mat_df_bin[i,2:ncol(mat_df_bin)])
}
# Fusion matrix and degrees
deg <-  full_join(deg_sender, deg_receiver, by="strain")
deg_temp <- deg
names(deg_temp)[names(deg_temp)=="strain"] <- "from"
mat_inhib_deg <- full_join(mat,deg_temp, by="from")
rm(deg_temp)
# Create a column -+ for Listeria strainsn threshold 250. 
mat_inhib_deg2 <- full_join(mat_inhib_deg,EGDe_inhib, by="from")
```

# INHIBITION MATRIX
```{r}
# Change the name of "strain" column to "from" in order to be able to make subsequent fusions with merge 
names(deg_receiver)[names(deg_receiver)=="strain"] <- "from"
names(deg_sender)[names(deg_sender)=="strain"] <- "from"
# merge 
send_ord <- mat %>%
  merge(deg_sender, on="from") %>%
  arrange(desc(Sender))

from <- send_ord$from

send_ord_sel <- select(send_ord, c(-1,-ncol(send_ord)))
colnames_send_ord <- names(send_ord_sel)
send_ord_sel_t <- t(send_ord_sel)
send_ord_sel_t <- as.data.frame(send_ord_sel_t)
names(send_ord_sel_t) <- from

send_ord_sel_t <- cbind(colnames_send_ord, send_ord_sel_t)

names(send_ord_sel_t)[names(send_ord_sel_t) == "colnames_send_ord"] <- "from"

send_receiv_ord <- send_ord_sel_t %>%
  merge(deg_receiver, on="from") %>%
  arrange(desc(Receiver))

rownames(send_receiv_ord)<-send_receiv_ord$from
send_receiv_ord <- select(send_receiv_ord, c(-ncol(send_receiv_ord),-1))

# Transpose the dataframe in order to have senders=rows  receivers=columns
send_receiv_ord_t <-t(send_receiv_ord)

```

```{r fig.width=8, fig.height=8}
hist(send_receiv_ord_t, col="blue", main ="GII", xlab = "")
```

# NEW STRAINS ADDED IN THE INHIBITION GRAPH
In Ramia et al., 2020, we described the antagonistic properties of 73 strains. 
In this study, we have added the data of 3 additionnal strains :
```{r}
a <- as.character(from)
b <- as.vector(unlist(strains_ramia))
setdiff(a,b)
rm(a,b)
```

# NESTEDNESS STRUCTURE
```{r fig.width=17, fig.height=17}
# Heatmap
send_receiv_ord <- as.matrix(send_receiv_ord)

# prepare de data for left and top annotation in order to add the genome availibility 
## vector contaning the order of appearance of strains in the ordered heatmap :
send_vec <- colnames(send_receiv_ord)
receiv_vec <- rownames(send_receiv_ord)
## order the genome dataframe according to send_vec and receiv_vec
genome_send <- genome %>% 
  arrange(factor(strain, levels=send_vec))
genome_receiv <- genome %>% 
  arrange(factor(strain, levels=receiv_vec))
## Define the parameters which will be used in the heatmap function for left_annotation and top_annotation
ha_send_left = rowAnnotation(genome = genome_send$genome,
                             col = list(genome = c("0" = "white", 
                                                   "1" = "grey25")),
                             gp = gpar(col = "black"),
                             show_legend = FALSE#,
                             #show_annotation_name = FALSE
                             )

ha_receiv_top = HeatmapAnnotation(genome  = genome_receiv$genome,
                                  height = unit(3, "cm"),
                                  col=list(genome = c("0" = "white",
                                                      "1" = "grey25")),
                                  gp = gpar(col = "black"),
                                  show_legend = FALSE
)

# Draw the heatmap
p <- Heatmap(send_receiv_ord_t,
        name = "Inhibition\nintensity", #title of legend
        show_heatmap_legend = FALSE,
        column_title = "Receiver", row_title = "Sender",
        column_title_gp = gpar(fontsize = 20),
        row_title_gp = gpar(fontsize = 20),
        row_names_gp = gpar(fontsize = 15),# Text size for row names
        column_names_gp = gpar(fontsize = 15), # Text size for column names
        row_names_side = "left", # names of the rows on the left
        column_names_side = "top",
        column_dend_side = "top", # dendrogram on the left
        row_dend_width = unit(4, "cm"), # size of the dendrogram on rows
        column_dend_height = unit(4, "cm"),# size of the dendrogram on columns
        col = inferno(100),
        cluster_columns=FALSE,
        cluster_rows = FALSE,
        top_annotation = ha_receiv_top,
        left_annotation = ha_send_left
)

p
#Export
tiff(file.path("output","nestedness.tiff"),
     units="in", width=17,height=17,res=1000)
p
dev.off()
rm(p)
```


# WEIGHT OF INHIBITION = f(Sender)
## Graphics
```{r fig.cap="only the strains for which the genome was sequenced and the sender degree is >0 are labeled with the strain name"}
r <- mat_inhib_deg2
mat_0_NA <- mat
mat_0_NA <- mat_0_NA %>% 
  select("from":"F2") %>% 
  pivot_longer(cols="ATCC35586":"F2",
               names_to="to",
               values_to="weight") %>% 
  mutate(weight=ifelse(weight<300,0,weight)) %>% 
  filter(weight!=0) %>% 
  group_by(from)%>%
  summarize(mean_weight= mean(weight, na.rm=TRUE))

t <- full_join(r,mat_0_NA,by="from")

t$EGDe_bin <- as.factor(t$EGDe_bin)
p <- ggplot(t, aes(x=Sender, y=mean_weight))+
  geom_smooth(method = "lm", 
              color="snow4", 
              fill="snow2")+
  geom_point(alpha = 0.8, aes(color=EGDe_bin))+
  scale_color_manual(values=c("purple4","red3"))+
  geom_label_repel (aes(label = ifelse(from %in% genome_yes & Sender>0, from, ""), color=EGDe_bin), # label only the strains for which Sender degree >20
                  max.overlaps = 100,
                  size= 2.8,
                  segment.color = ifelse(mat_inhib_deg2$EGDe_bin=="0",
                                         "purple4", 
                                         "red3"),
                  min.segment.length = 0,
                  box.padding = 0.5,
                  label.padding = 0.15,
                  label.r = 0.13,
                  label.size =0.4
                  )+
  labs(x="Sender degree", y="Mean inhibition weight (min)")+
  theme_base()+
  theme(legend.position = "none")
p
#Export
tiff(file.path("output","weight-send.tiff"),
     units="in", width=6,height=6,res=1000)
p
dev.off()
rm(p)
```

## Linear regression
Model
```{r}
cor.test(t$Sender,t$mean_weight, method=c("pearson"))
```
p-value <0.05, the null hypothesis is rejected. 

```{r}
cor(t$Sender,t$mean_weight, method=c("pearson"))

```
the correlation coeficient is close to 1

```{r}
lm1 <- lm(t$Sender~t$mean_weight)
lm1
```


```{r}
summary(lm1)
```


Are the residuals normality distributed ?
```{r}
hist(lm1$residuals)
```
the distribution of the data are close to normal distribution.

Autocorrelation test
```{r}
library(car)
durbinWatsonTest(lm1)
```
The null hypothesis is not rejected, the residuals are independant, there is no autocorrelation of the data. 



# CLUSTERING HEATMAP
```{r fig.width=15, fig.height=9}
mat_inhib_deg2_genome <- mat_inhib_deg2 %>% 
  filter(from %in% genome_yes)
rownames(mat_inhib_deg2_genome) <- mat_inhib_deg2_genome$from
mat_inhib_deg2_genome <- mat_inhib_deg2_genome[,-1]


row_ha <- rowAnnotation("anti-Listeria"=anno_barplot(mat_inhib_deg2_genome$EGDe,
                                                     width = unit(4,"cm")))
mat_inhib_deg2_genome_1_76 <- select(mat_inhib_deg2_genome, 1:76)

for (i in 1:ncol(mat_inhib_deg2_genome_1_76)) {   
  mat_inhib_deg2_genome_1_76[ , i]=ifelse(mat_inhib_deg2_genome_1_76[ , i]>300, 1, 0)
}

p <- Heatmap(mat_inhib_deg2_genome_1_76,
        cluster_columns = FALSE,
        #name = "Inhibition\nintensity", #title of legend
        show_heatmap_legend = FALSE,
        column_title = "Receiver", row_title = "Sender",
        column_title_gp = gpar(fontsize = 20),
        row_title_gp = gpar(fontsize = 20),
        row_names_gp = gpar(fontsize = 18),# Text size for row names
        column_names_gp = gpar(fontsize = 10), # Text size for column names
        row_names_side = "left", # names of the rows on the left
        column_names_side = "top",
        column_dend_side = "top", # dendrogram on the left
        row_dend_width = unit(4, "cm"), # size of the dendrogram on rows
        col = viridis(100),
        clustering_distance_rows ="euclidean",
        clustering_method_rows = "average",
        left_annotation = rowAnnotation(foo = anno_block(gp = gpar(fill = 0:0),
        labels = c("gp1", "gp2", "gp3"), 
        labels_gp = gpar(col = "black", fontsize = 15))),
        row_km = 3,
        right_annotation = row_ha
)
p
#Export
tiff(file.path("output","clust_heatmap.tiff"),
     units="in", width=17,height=10,res=1000)
p
dev.off()
rm(p)
```

# GROUPS GP1, GP2, GP3: STATISTICAL ANALYSIS
```{r}
# add gp1,gp2 and gp3 as categorical variable
strains <- c("10040100629",
             "F2",
             "F88",
             "9.4",
             "DSM20344")
  
gp <- "gp1"
gp1 <- cbind(strains, gp)

strains <- c("F4",
             "IFIP 710",
             "F84",
             "F14",
             "CIP100481",
             "F7",
             "CP5",
             "CP4",
             "LMA28",
             "CIP102035",
             "JIP 28/91",
             "CP1",
             "CP14",
             "8.1",
             "F42",
             "MM 3364.01",
             "MM 3365.01")
gp <- "gp2"
gp2 <- cbind(strains,gp)


strains <- c("F48",
             "F73",
             "CIP101354",
             "LLS R 919",
             "DSM20590",
             "JIP 05/93",
             "RFA 378") 
gp <- "gp3"
gp3 <- cbind(strains,gp)


gp <- rbind(gp1,gp2,gp3)
mat_inhib_deg2_genome <- rownames_to_column(mat_inhib_deg2_genome, var = "strains")

rm(strains)

mat_inhib_deg2_genome <- as.data.frame(mat_inhib_deg2_genome)
gp <- as.data.frame(gp)
mat_inhib_deg2_genome <- full_join(mat_inhib_deg2_genome, gp, by="strains")
```

```{r}
lm1 <- lm(EGDe~gp, data = mat_inhib_deg2_genome)
summary(lm1)
anova(lm1)
aov1 <- aov(EGDe~gp, data = mat_inhib_deg2_genome)
anova(aov1)
```

```{r}
tuk <- TukeyHSD(aov1)
tuk <- as.data.frame(tuk$gp)
tuk <- rownames_to_column(tuk, var="gp")
tuk <- tuk %>% 
  separate(col= gp,
           into = c("group1", "group2"),
           sep = "-") %>% 
  mutate(p_vals=formatC(`p adj`, format = "e", digits = 2)) %>% 
  select(group1,group2, `p_vals`) 

y.position <- c(800,750,200)
tuk <- cbind(tuk, y.position)
```


```{r}
base <- ggplot(mat_inhib_deg2_genome, aes(x = gp, y = EGDe)) +
  geom_boxplot(notch = FALSE,
               color="mistyrose4",
               fill="mistyrose4",
               alpha=0.3
               ) +
  geom_jitter(width = 0.1,
              height = 0.1,
              alpha = 0.3) +
  scale_y_continuous(limits = c(0, 850)) +
  labs(x = "", y = "Inhibition weight (min)") +
  theme_base()


p <- base + 
  add_pvalue(tuk, label = "p = {`p_vals`}", tip.length = 0, label.size = 4)

p

#Export
tiff(file.path("output","anov_weight.tiff"),
     units="in", width=3,height=7,res=1000)
p
dev.off()
rm(p)
```
```{r}
rm(tuk, base, y.position)
```

# OCCURENCE
## DATA PREPARATION
```{r}
occ <-
  read.csv2(
    here::here("data-clean", "occurence3_no_imm2.csv"),
    check.names = F,
    header = T,
    stringsAsFactors = T
  )

occ <- occ %>% 
  arrange(factor(Strains, levels = gp$strains))

bacteriocin_new <-
  read.csv2(
    here::here("data-clean", "bacteriocin_new.csv"),
    check.names = F,
    header = T,
    stringsAsFactors = T
  )
bacteriocin_new$new <- as.factor(bacteriocin_new$new)
```

```{r}
mat_occ <- as.matrix(occ)
rownames(mat_occ) <- mat_occ[,1] 
mat_occ <- mat_occ[,-1]
```

## HEATMAP
```{r}
# calculate the meanweight of inhibition of receivers for each strain for which the genome was sequenced 
weight <- mat
weight <- weight %>% 
select("from":"F2") %>% 
  pivot_longer(cols="ATCC35586":"F2",
               names_to="to",
               values_to="weight") %>% 
  mutate(weight=ifelse(weight<300,NA,weight)) %>% 
  filter(from %in% genome_yes) %>% 
  group_by(from)%>%
  summarize(mean_weight= mean(weight, na.rm=TRUE))
weight$from <- droplevels(weight$from)
weight <- weight[match(rownames(mat_occ), weight$from),]
 

# define the colors 
colors = structure(
  c("gray93","paleturquoise3", "palevioletred3"),#"seashell"
  names = c(0,1,2)
)
t <-mat_inhib_deg2_genome %>% 
  select(strains,EGDe_bin)

colnames(t)[colnames(t) == 'EGDe_bin'] <- 'EGDe'

gp <- left_join(gp,t, by="strains")

ha_EGDe <-  rowAnnotation(EGDe = gp$EGDe,
                          col = list(EGDe = c("0" = "white", 
                                              "1" = "gray23")),
                          gp = gpar(col = "black"),
                          show_legend = FALSE#,
                             #show_annotation_name = FALSE
                             )

p <- Heatmap(
  mat_occ,
  
  # NO LEGENDS 
  show_heatmap_legend = FALSE,
  
  # TITLES
  column_title = "Bacteriocin",
  row_title = "Strain",
  
  # Add frame to each cells
  rect_gp = gpar(col = "white", lwd = 1),
  # Text size for row names
  row_names_gp = gpar(fontsize = 9),
  # Text size for column names
  column_names_gp = gpar(fontsize = 9),
  
  # names of the rows on the left
  row_names_side = "left",
  column_names_side = "top",
  col = colors,
  
  
  # SPLIT COLUMNS
  column_split = c(
    rep("class I", 4),
    rep("Class IIa", 7),
    rep("class IIc", 1)
  ),
  column_gap = unit(3, "mm"),
  
  
  # TOP ANNOTATION
  top_annotation = HeatmapAnnotation(
    new  = bacteriocin_new$new,
    col=list(new = c("0" = "white", "1" = "plum3")), #"seagreen3"
    simple_anno_size = unit(0.3,"cm"),
    show_legend = FALSE,
    show_annotation_name = FALSE,
    foo = anno_block(
      gp = gpar(fill = 0:0),
      height = unit(0.55, "cm"),
      labels = c("I", "IIa", "IIc"),
      labels_gp = gpar(col = "black", 
                       fontsize = 10),
      )
   
    ),
  column_km = 3,
  row_split = gp$gp,
  
  # LEFT ANNOTATION 
  left_annotation = rowAnnotation(foo = anno_block(gp = gpar(fill = 0:0),
                                                   labels = c("gp1", "gp2", "gp3"),
                                                   width = unit(0.55, "cm"),
                                                   labels_gp = gpar(col = "black", fontsize = 10)),
                                  "Mean inhibition\nweight"=anno_barplot(weight$mean_weight,
                                                                           width = unit(1,"cm"))),
  
  row_km = 3,
  
  # RIGHT ANNOTATION
  right_annotation = ha_EGDe
)
p

# Export
tiff(file.path("output","occurence.tiff"),
     units="in", width=6,height=7,res=1000)
p
dev.off()
rm(p)
```

# REFERENCES
```{r}
sessionInfo()
```


